DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car

### 2018-04-16 paper/background.tex
intro - overview revision

Using deep neural networks involves two distinct
phases~\cite{NVIDIA2015}. The first
phase is \emph{training} during which the weights of the network are
incrementally updated by backpropagating errors it sees from the
training examples. Once the network is trained---i.e., the weights of
the network minimize errors in the training examples---the next phase
is \emph{inferencing}, during which unseen data is fed to the network
as input to produce predicted output (e.g., predicted image
classification). In general, the training phase is more computationally
intensive and requires high throughput, which is generally not
available on embedded platforms. The inferencing phase, on the
other hand, is relatively less computationally intensive and latency becomes
as important, if not moreso, as computational throughput, because many
use cases have strict real-time requirements (e.g., search query
latency).

@@

\subsection{Embedded Computing Platforms for Real-Time Inferencing}
%% \fixme{we can consider moving this section to a dedicated related work
%%   section at the end.}
Real-time embedded systems, such as an autonomous vehicle, present
unique challenges for deep learning, as the computing platforms of such
systems must satisfy two often conflicting goals~\cite{Otterness2017}:
%% Recent successes in AI, including NVIDIA's DAVE-2 showing, are due
%% in large part to the increased computing performance,
%% which afforded researchers to train and use ever deeper neural networks with
%% high accuracy.
%% For practical applications, the computer platform in an
%% autonomous vehicle must satisfy two often conflicting goals:
(1) The platform must provide 
enough computing capacity for real-time processing of computationally
expensive AI workloads (deep neural networks); and
(2) The platform must also satisfy various
constraints such as cost, size, weight, and power consumption limits.

@@

The \emph{primary objectives of this study} are to understand (1) the
necessary computing performance for applying AI technology-based
robotics systems, and (2) what kind of computing architecture and
runtime supports are most appropriate for such a workload. To
achieve these goals, we implement a low-cost autonomous car platform
as a case-study and systematically conduct experiments, which we will 
describe in the subsequent sections.

>>>>

Using deep neural networks involves two distinct
phases~\cite{NVIDIA2015}. The first
phase is \emph{training} during which the weights of the network are
incrementally updated by backpropagating errors it sees from the
training examples. Once the network is trained---i.e., the weights of
the network minimize errors in the training examples---the next phase
is \emph{inferencing}, during which unseen data is fed to the network
as input to produce predicted output (e.g., predicted image
classification). In general, the training phase is more computationally
intensive and requires high throughput, which is generally not
available on embedded platforms. The inferencing phase, on the
other hand, is relatively less computationally intensive and latency becomes
as important, if not moreso, as computational throughput, because many
use cases have strict real-time requirements.

%% (e.g., search query latency)

@@

\subsection{Embedded Computing Platforms for Real-Time Inferencing}
%% \fixme{we can consider moving this section to a dedicated related work
%%   section at the end.}
Real-time embedded systems, such as an autonomous vehicle, present
unique challenges for deep learning, as the computing platforms of such
systems must satisfy two often conflicting goals:
%% Recent successes in AI, including NVIDIA's DAVE-2 showing, are due
%% in large part to the increased computing performance,
%% which afforded researchers to train and use ever deeper neural networks with
%% high accuracy.
%% For practical applications, the computer platform in an
%% autonomous vehicle must satisfy two often conflicting goals:
(1) The platform must provide 
enough computing capacity for real-time processing of computationally
expensive AI workloads (deep neural networks); and
(2) The platform must also satisfy various
constraints such as cost, size, weight, and power consumption limits~\cite{Otterness2017}.

@@

The \emph{primary objectives of this study} are to understand (1) the
necessary computing performance for applying AI technology-based
robotics systems, and (2) the characteristics of the computing
platform to support such a workload.
To achieve these goals, we implement a low-cost autonomous car platform
as a case-study and systematically conduct experiments, which we will 
describe in the subsequent sections.

### 2018-04-16 paper/evaluation.tex
intro - overview revision

The Raspberry Pi 3 Model B platform used in DeepPicar equips a Broadcom
BCM2837 SoC, which has a quad-core ARM Cortex-A53 cluster,
running at up to 1.2GHz. Each core has a 32K private Instruction cache
and a 32K private Data cache, and all cores share a 512KB L2 cache.
The chip also includes Broadcom's Videocore IV
GPU, although we did not use the GPU in our evaluation due to the lack
of sofware support (TensorFlow is not compatible with the Raspberry Pi's GPU).
For software, we use Ubuntu MATE 16.04, TensorFlow 1.1 and Python
2.7. We disabled DVFS (dynamic voltage frequency scaling) and
configured the clock speed of each core statically at the maximum 1.2GHz.

@@

Second, we find that the measured average
execution time of a single control loop is 22.87 ms, or 43.7 Hz and
the 99 percentile time is 23.38 ms.
This means that the DeepPicar can operate
at up to 40 Hz control frequency in real-time using only the on-board
Raspberry Pi 3 computing platform, as no remote computing resources 
were necessary. We consider these results surprising given the complexity
of the deep neural network, and the fact that the inference operation
performed by TensorFlow only utilizes the CPU cores of the
Raspberry Pi 3 (its GPU is not supported by Tensorflow).
In comparison, NVIDIA's DAVE-2 system, which has the exact same neural
network architecture, reportedly runs at 30 Hz~\cite{Bojarski2016}. 
Although we believe it was not
limited by their computing platform (we will experimentally compare
performance differences among multiple embedded computing platforms,
including NVIDIA's Jetson TX2, later in
Section~\ref{sec:comparison}), the fact that a low-cost
Raspberry Pi 3 can achieve comparable real-time control performance is
surprising.

>>>>

The Raspberry Pi 3 Model B platform used in DeepPicar equips a Broadcom
BCM2837 SoC, which has a quad-core ARM Cortex-A53 cluster,
running at up to 1.2GHz. Each core has a 32K private Instruction cache
and a 32K private Data cache, and all cores share a 512KB L2 cache.
The chip also includes Broadcom's Videocore IV GPU, although we did
not use the GPU in our evaluation, due to the lack of sofware support
\footnote{TensorFlow currently only support NVIDIA's GPUs.}.
For software, we use Ubuntu MATE 16.04, TensorFlow 1.1.
We disabled DVFS (dynamic voltage frequency scaling) and
configured the clock speed of each core statically at the maximum 1.2GHz.

@@

Second, we find that the measured average
execution time of a single control loop is 22.87 ms, or 43.7 Hz and
the 99 percentile time is 23.38 ms.
This means that the DeepPicar can operate
at up to 40 Hz control frequency in real-time using only the on-board
Raspberry Pi 3 computing platform, as no remote computing resources 
were necessary. We consider these results surprising given the complexity
of the deep neural network, and the fact that the inference operation
performed by TensorFlow only utilizes the CPU cores of the Raspberry Pi 3.
In comparison, NVIDIA's DAVE-2 system, which has the exact same neural
network architecture, reportedly runs at 30 Hz~\cite{Bojarski2016}. 
Although we believe it was not
limited by their computing platform (we will experimentally compare
performance differences among multiple embedded computing platforms,
including NVIDIA's Jetson TX2, later in
Section~\ref{sec:comparison}), the fact that a low-cost
Raspberry Pi 3 can achieve comparable real-time control performance is
surprising.

### 2018-04-16 paper/intro.tex
Updates to paper
intro - overview revision

% related work and remaining problems
To understand what kind of computing hardware is needed for AI
workloads, we need a testbed and realistic workloads. While using a real 
car-based testbed would be most ideal, it is not only highly expensive, but also
poses serious safety concerns that hinder development and exploration.
Therefore, there is a strong need for safer and less costly
testbeds. There are already several relatively inexpensive RC-car
based testbeds, such as MIT's 
RaceCar~\cite{shin2017project} and UPenn's F1$/$10 racecar~\cite{upennf1tenth}.
However, these RC-car testbeds still cost more than \$3,000, requiring
considerable investment.

% our goals
Instead, we want to build a low cost testbed that still employs the
state-of-the art AI technologies. Specifically, we focus on an end-to-end
deep learning based real-time control system,
which was developed for a real self-driving car, NVIDIA
DAVE-2~\cite{Bojarski2016}, and uses the same methodology on a
smaller and less costly setup. In developing the testbed, our
goals are (1) to analyze real-time issues in DNN based end-to-end
control; and (2) to evaluate real-time performance of contemporary embedded
platforms for such a workload.

% DeepPicar introduction
In this paper, we present DeepPicar, a low-cost autonomous car
platform for research. From a hardware perspective,
DeepPicar is comprised of a Raspberry Pi 3 Model B quad-core
computer, a web camera and a RC car, all of which are affordable
components (less than \$100 in total).
The DeepPicar, however, employs state-of-the-art AI
technologies, including a vision-based end-to-end control system that
utilizes a deep convolutional neural network (CNN).
The network receives an image frame from a single forward
looking camera as input and generates a predicted steering angle
value as output at each control period in \emph{real-time}.
The network has 9 layers, about 27 million connections
and 250 thousand parameters (weights).
The network architecture is almost identical to that of NVIDIA's DAVE-2
self-driving car~\cite{Bojarski2016}, which uses a significantly more powerful
computer (Drive PX computer~\cite{drivepx}) than a Raspberry Pi~3.
We chose to use a Pi 3 not only because it is affordable, but also because it is representative
of today's mainstream low-end embedded multicore platforms found in
smartphones and other embedded devices.

@@

% how we trained (maybe moved to a later section)
We apply a standard imitation learning methodology to train the car to
follow tracks on the ground. We collect data for
training and validation by manually
controlling the RC car and recording the vision (from the webcam
mounted on the RC-car) and the human control inputs. We then train the
network offline using the collected data on a desktop computer, which
is equipped with a NVIDIA GTX 1060 GPU. Finally, the trained network is copied
back to the Raspberry Pi 3, which is then used to perform inference
operations---locally on the Pi 3---in the car's main control loop in
real-time. For real-time control, each inference operation must
be completed within the desired control period (e.g., 33.$\overline{\mbox{33}}$ ms
 period for 30 Hz control frequency).
% how we evaluated (in terms real-time performance)

% what are our findings?
Using the DeepPicar platform, we systematically analyze its real-time
capabilities in the context of deep-learning based real-time
control, especially on real-time deep neural network inferencing.
We also evaluated other, more powerful, embedded computing
platforms to better understand achievable real-time performance of
DeepPicar's deep-learning based control system and the impact of
computing hardware architectures.

% Findings:
%
% - we find DNN processing is an ideal real-time workload from the
%   perspective of timing predictability as the amount of computation
%   needed is fixed at design time and doesn't change over input.
%
% - real-time processing of non-trivial DNN inferencing is feasible
%   on today's embedded multicore platform, even as inexpensive as
%   raspberry pi3.
%
% - all tested platforms have enough computing capacity to consolidate 
%   multiple DeepPicar's CNN workloads or other compute intensive tasks
%   simultaneously.
%
% - however, when consolidating multiple workloads shared resource
%   contention is an important issue. we carefully analyzed
%   sensitivity of DeepPicar's workload with respect to shared
%   hardware resources.
%
% - in particular, we find DNN processing is highly sensitive to
%   memory performance. performance severely affected when memory
%   bandwidth is contented.
%
% - however, we find that dnn processing is not sensitive to cache
%   space. we find applying cache partitioning does not help improve
%   isolation of DNN workload when memory bandwidth is contended.
From the systematic study, we have made a number of interesting
observations from the perspective of real-time systems.
% predictability
First, we find that DNN inferencing is highly predictable---from the
timing perspective---as the amount of computation needed to complete a
single inference is fixed at the DNN architecture design time and does
not change at runtime over different inputs (e.g., different image
frames). This predictable timing behavior is obviously a desirable
property for real-time systems.
%% , which gives another reason for the
%% real-time community to be interested in the use of DNNs.

% performance on low-cost embedded multicore
Second, we find that real-time processing of DeepPicar's CNN is
feasible on today's embedded computing platforms, even one  
as inexpensive as the Raspberry Pi 3. We find that the control loop 
completes in under 33.$\overline{\mbox{33}}$ ms, or 30 Hz, using just two 
cores of the Raspberry Pi 3's quad-core CPU (w/o using its GPU), and can
do so 100\% of the time; or we can achieve 20 Hz control frequency using
just one core.
Other tested embedded platforms, the Intel UP and NVIDIA TX2, offer even
better performance, and are capable of supporting deep-learning based 
real-time control over 100 Hz control frequency on the TX2 when its
embedded GPU is used.
We find that all the tested embedded computing platforms have 
enough computing capacity to consolidate multiple instances of
DeepPicar's DNN workload.

% all platforms have enough capacity to consolidate
Third, contention in shared hardware resources remains an important
issue when consolidating multiple workloads. 
In particular, we find that DNN inferencing is highly sensitive to
memory performance interference, as we observe up to an 11.6X control
loop execution time increase when memory performance intensive
applications are co-scheduled on idle CPU cores.
On the other hand, we find that DNN inferencing is not sensitive to
cache space.
%% ; we show that applying cache partitioning has negligible impact
%% in providing timing isolation on consolidated scenarios.
This finding
suggests that in order to guarantee real-time performance of DNN-based
control applications, solutions to guarantee their memory performance
are more important.

% contributions
The {\bf contributions} of this paper are as follows:
\begin{itemize}
  \item We present DeepPicar, a low-cost autonomous car testbed, which
    is based on a state-of-the-art end-to-end deep learning, as 
    open-source~\footnote{The source code, and build 
instructions for the DeepPicar
can be found at: \url{https://github.com/heechul/picar}}, with the
    goal of lowering the economic and safety-related barriers to the study
    of autonomous cars.

  \item We systematically analyze real-time characteristics of DeepPicar's
    DNN inferencing workload on a number of representative
    contemporary embedded computing platforms. While larger embedded
    platforms have been evaluated, as in  \cite{Otterness2017}, to the
    best of our knowledge, no prior work investigated the
    impacts of shared resource contention on DNN inferencing workloads
    in the context of smaller embedded platforms like the Raspberry Pi 3.

\end{itemize}

>>>>

% related work and remaining problems
To understand what kind of computing hardware is needed for AI
workloads, we need a testbed and realistic workloads. While using a
real car-based testbed would be most ideal, it is not only highly
expensive, but also poses serious safety concerns that hinder
development and exploration. Therefore, there is a need for safer and
less costly testbeds.

%% There are already several relatively inexpensive RC-car
%% based testbeds, such as MIT's 
%% RaceCar~\cite{shin2017project} and UPenn's F1$/$10 racecar~\cite{upennf1tenth}.
%% However, these RC-car testbeds still cost more than \$3,000, requiring
%% considerable investment.

%% % our goals
%% Instead, we want to build a low cost testbed that still employs the
%% state-of-the art AI technologies. Specifically, we focus on an end-to-end
%% deep learning based real-time control system,
%% which was developed for a real self-driving car, NVIDIA
%% DAVE-2~\cite{Bojarski2016}, and uses the same methodology on a
%% smaller and less costly setup. In developing the testbed, our
%% goals are (1) to analyze real-time issues in DNN based end-to-end
%% control; and (2) to evaluate real-time performance of contemporary embedded
%% platforms for such a workload.

% DeepPicar introduction
In this paper, we present DeepPicar, a low-cost autonomous car
testbed for research. From a hardware perspective,
DeepPicar is comprised of a Raspberry Pi 3 Model B quad-core
computer, a web camera and a RC car, all of which are affordable
components (less than \$100 in total).
The DeepPicar, however, employs a state-of-the-art AI
technology, namely deep learning based end-to-end real-time control,
which utilizes a deep convolutional neural network (CNN).
The CNN receives an image frame from a single forward
looking camera as input and generates a predicted steering angle
value as output at each control period in \emph{real-time}.
The CNN has 9 layers, about 27 million connections
and 250 thousand parameters (weights).
DeepPicar's CNN architecture is identical to that of NVIDIA's
real-sized self-driving car, called DAVE-2~\cite{Bojarski2016}, which
drove on public roads without human driver's intervention while only
using the CNN, which runs on a significantly more powerful computing
platform (Drive PX~\cite{drivepx}).

%% , which uses a significantly more
%% powerful computer (Drive PX computer~\cite{drivepx}) than a Raspberry
%% Pi~3.
%% We chose to use a Pi 3 not only because it is affordable, but 
%% also because it is representative of today's mainstream low-end
%% embedded multicore platforms found in smartphones and other embedded
%% devices.

@@

% how we trained (maybe moved to a later section)
%% We apply a standard imitation learning methodology to train the car to
%% follow tracks on the ground. We collect data for
%% training and validation by manually
%% controlling the RC car and recording the vision (from the webcam
%% mounted on the RC-car) and the human control inputs. We then train the
%% network offline using the collected data on a desktop computer, which
%% is equipped with a NVIDIA GTX 1050 Ti GPU. Finally, the trained network is copied
%% back to the Raspberry Pi 3, which is then used to perform inference
%% operations---locally on the Pi 3---in the car's main control loop in
%% real-time. For real-time control, each inference operation must
%% be completed within the desired control period (e.g., 33.$\overline{\mbox{33}}$ ms
%%  period for 30 Hz control frequency).
% how we evaluated (in terms real-time performance)

% what are our findings?
Using DeepPicar, we systematically analyze its real-time
capabilities in the context of deep-learning based real-time
control, especially on real-time deep neural network \emph{inferencing}.
We also evaluated other, more powerful, embedded computing
platforms to better understand achievable real-time performance of
DeepPicar's deep-learning based control system and the impact of
computing hardware architectures.

% Findings:
%
% - we find DNN processing is an ideal real-time workload from the
%   perspective of timing predictability as the amount of computation
%   needed is fixed at design time and doesn't change over input.
%
% - real-time processing of non-trivial DNN inferencing is feasible
%   on today's embedded multicore platform, even as inexpensive as
%   raspberry pi3.
%
% - all tested platforms have enough computing capacity to consolidate 
%   multiple DeepPicar's CNN workloads or other compute intensive tasks
%   simultaneously.
%
% - however, when consolidating multiple workloads shared resource
%   contention is an important issue. we carefully analyzed
%   sensitivity of DeepPicar's workload with respect to shared
%   hardware resources.
%
% - in particular, we find DNN processing is highly sensitive to
%   memory performance. performance severely affected when memory
%   bandwidth is contented.
%
% - however, we find that dnn processing is not sensitive to cache
%   space. we find applying cache partitioning does not help improve
%   isolation of DNN workload when memory bandwidth is contended.
From the systematic study, we have made a number of interesting
observations.
% predictability
First, we find that CNN inferencing is highly time predictable as the
amount of computation needed to complete a 
single inference is fixed at the network architecture design time and
does not change at runtime over different inputs (e.g., different image
frames). This predictable timing behavior is obviously a desirable
property for real-time systems.
%% , which gives another reason for the
%% real-time community to be interested in the use of DNNs.

% performance on low-cost embedded multicore
Second, we find that real-time processing of CNN inferencing is
feasible on today's embedded computing platforms, even one   
as inexpensive as the Raspberry Pi 3. For example, just one Cortex-A53
core of the Raspberry Pi3 is able to complete the CNN inferencing task
in less than 50ms, sufficient to support 20 Hz control loop. 
%% We find that the control loop 
%% completes in under 33.$\overline{\mbox{33}}$ ms, or 30 Hz, using just two 
%% cores of the Raspberry Pi 3's quad-core CPU (w/o using its GPU), and can
%% do so 100\% of the time; or we can achieve 20 Hz control frequency using
%% just one core.
%% Other tested embedded platforms, the Intel UP and NVIDIA TX2, offer even
%% better performance, and are capable of supporting deep-learning based 
%% real-time control over 100 Hz control frequency on the TX2 when its
%% embedded GPU is used.
We find that all the tested embedded computing platforms have 
enough computing capacity to consolidate multiple instances of
DeepPicar's DNN workload.

% all platforms have enough capacity to consolidate
Third, contention in shared hardware resources remains an important
issue when consolidating multiple workloads. 
We observe up to 11.6X control
loop execution time increase due to shared resource contention when
memory intensive  applications are co-scheduled on idle CPU cores.
We find that DeepPicar's inferencing task is highly
sensitive to memory bandwidth, but not cache space.

%% ; we show that applying cache partitioning has negligible impact
%% in providing timing isolation on consolidated scenarios.
%% This finding
%% suggests that in order to guarantee real-time performance of DNN-based
%% control applications, solutions to guarantee their memory performance
%% are more important.

% contributions
The {\bf contributions} of this paper are as follows:
\begin{itemize}
  \item We present DeepPicar, a low-cost autonomous car testbed using
    a state-of-the-art CNN for end-to-end real-time control, to lower
    economic and safety-related barriers to the study autonomous
    cars.~\footnote{The source code, dataset, and build instructions
      of DeepPicar can be found at:\fixme{Michael github url}}

  \item We provide extensive systematic experimental results using
    DeepPicar's CNN inferencing workload on a number of representative
    contemporary embedded computing platforms. The results provide
    insights for real-time processing of the CNN workload and the
    effect of consolidation.

    %% While larger embedded
    %% platforms have been evaluated, as in  \cite{Otterness2017}, to the
    %% best of our knowledge, no prior work investigated the
    %% impacts of shared resource contention on DNN inferencing workloads
    %% in the context of smaller embedded platforms like the Raspberry Pi 3.
\end{itemize}

### 2018-04-16 paper/overview.tex
intro - overview revision

For the neural network architecture, we implement
NVIDIA DAVE-2's convolutional neural network (CNN) using a TensorFlow
based CNN model in ~\cite{deeptesla}. Note, however, that the CNN model
in~\cite{deeptesla} was considerably larger than NVIDIA's CNN model as
it contained an additional fully-connected layer of approximately 1.3M
additional parameters. We eliminated the additional layer to
faithfully recreate the NVIDIA's original CNN model.
%% The main difference is that we do not utilize a normalization layer, and 
%% instead initialize the weights using a Xavier initialization. 
As in DAVE-2, the CNN takes a raw color image (200x66 RGB pixels)
as input and produces a single steering angle value as
output. Figure~\ref{fig:architecture} shows the network architecture
used in this paper, which is comprised of 9 layers, 250K parameters,
and about 27 million connections as in NVIDIA DAVE-2's. Note, however,
that we did not apply the normalization mentioned
in~\cite{Bojarski2016}, as it does not include trainable parameters
and its computational demand with respect to the overall CNN
processing is minimal.

@@

% data collection and training.
To collect the training data, a human pilot manually drives the RC car
on tracks we created (Figure~\ref{fig:tracks}) to record
timestamped videos and contol commands. The stored data is then copied 
to a desktop computer, which is equipped with a NVIDIA Titan Xp GPU, 
where we train the network to accelerate training speed. 
For comparison, training the network on the Raspbeery Pi 3 takes
approximately 4 hours, whereas it takes only about 4 minutes on the
desktop computer using the Titan Xp GPU.

@@

% inferencing on pi3
Once the network is trained on the desktop computer, the trained model
is copied back to the Raspberry Pi 3. The network is then used
by the car's main controller, which feeds an image frame from the web
camera as input to the network. In each control period, the produced 
steering angle output is then converted into the PWM values of the 
steering motor of the car. Figure~\ref{fig:controlloop} shows simplified 
pseudo code of the controller's main loop. Among the five steps, the 3rd step, 
network inferencing, is the most computationally intensive and dominates the
execution time.

Note that although the steering angle output of the network ($angle$) is
a continuous real value, the RC car platform we used unfortunately
only supports three discrete angles---left (-30$^{\circ}$), center
(0$^{\circ}$), and right (+30$^{\circ}$)---as control inputs.
Currently, we approximate the network generated real-valued
angle to the closest one of the three angles, which may
introduce inaccuracy in control.
In the future, we plan to use a different (more expensive) RC car
platform that can precisely control the car's steering angle.

Other issues we observed in training/testing the network, which
can affect the prediction accuracy of the CNN, are camera and actuator
(motor) control latencies. In DeepPicar's
context, the camera latency  measured from the time the camera
sensor observes the scene to the time the computer actually reads the
digitized image data. This time can be noticable depending on the camera 
used and the performance of the Pi. Higher camera latency could
negatively affect control performance, because the DNN would analyze
stale scenes. Actuator (motor) control latency also can be a
factor, because moving the steering motor to a desired position takes
time. 
In our platform, the combined latency is measured to be around
50 ms. In other words, in the recorded video and control data, a
control action (left, center, right) taken by the CNN appears to be
applied 50 ms later. Considering that camera's framerate is 30Hz
(33.$\overline{\mbox{33}}$ ms/frame), this is about two frames after the 
control action.
%% We experimentally measured the camera
%% latency and found it to be around 50-100 ms.
If this value is too high, control performance may suffer.
Our initial prototype suffered this problem as the observed camera and
control latency was as high as 300 ms, which negatively affected
control performance.
For reference, the latency of human perception is known to be as fast
as 13 ms~\cite{ThomasBurger2015}. 
% https://www.pubnub.com/blog/2015-02-09-how-fast-is-realtime-human-perception-and-technology/

Despite these issues, the trained models were still able to
achieve a reasonable degree of accuracy, successfully navigating
several different tracks we used for training. In our testing, we found 
that the DeepPicar could remain on a fairly challenging track for over 
10 minutes at a moderate speed (50\% throttle), at which point we stopped
the experiment, and more than one minute at a higher speed (75\%
throttle)~\footnote{A collection of self-driving videos can be found
  at: \url{https://photos.app.goo.gl/ce93sU7jPk4ywO8u2}}. Running at
higher speed is inherently more challenging because the CNN controller
has less time to recover from mistakes (bad predictions).  Also, we
find that the prediction accuracy is significantly affected by the
quality of training data as well as various environmental factors such
as lighting conditions. We plan to investigate more systematic ways
to improve the CNN's prediction accuracies.

We would like to stress, however, that the issues related to the
CNN's accuracies have no impact on the \emph{computational 
aspects of the system}, and that our main focus of this study is
not in improving network accuracy but in closely replicating the
DAVE-2's network architecture and studying its real-time
characteristics.

>>>>

For the neural network architecture, we implement
NVIDIA DAVE-2's convolutional neural network (CNN) using a TensorFlow
based CNN model in ~\cite{deeptesla}. Note, however, that the CNN model
in~\cite{deeptesla} was considerably larger than NVIDIA's CNN model as
it contained an additional fully-connected layer of approximately 1.3M
additional parameters. We eliminated the additional layer to
faithfully recreate the NVIDIA's original CNN model.
%% The main difference is that we do not utilize a normalization layer, and 
%% instead initialize the weights using a Xavier initialization. 
As in DAVE-2, the CNN takes a raw color image (200x66 RGB pixels)
as input and produces a single steering angle value as
output. Figure~\ref{fig:architecture} shows the network architecture
used in this paper, which is comprised of 9 layers, 250K parameters,
and about 27 million connections as in NVIDIA DAVE-2's.
%% Note, however,
%% that we did not apply the normalization mentioned
%% in~\cite{Bojarski2016}, as it does not include trainable parameters
%% and its computational demand with respect to the overall CNN
%% processing is minimal.

@@

% data collection and training.
To collect the training data, a human pilot manually drives the RC car
on tracks we created (Figure~\ref{fig:tracks}) to record
timestamped videos and contol commands. The stored data is then copied 
to a desktop computer, which is equipped with a NVIDIA Titan Xp GPU, 
where we train the network to accelerate training speed.

%% For comparison, training the network on the Raspbeery Pi 3 takes
%% approximately 4 hours, whereas it takes only about 4 minutes on the
%% desktop computer using the Titan Xp GPU.

@@

% inferencing on pi3
Once the network is trained on the desktop computer, the trained model
is copied back to the Raspberry Pi 3. The network is then used
by the car's main controlller, which feeds an image frame from the web
camera as input to the network at each control period. The network
produced steering angle output is then converted into the PWM values
of the car's steering motor. Figure~\ref{fig:controlloop} shows simplified 
pseudo code of the controller's main loop. Among the five steps, the 3rd step, 
network inferencing, is the most computationally intensive and dominates the
execution time.

Note that although the steering angle output of the network ($angle$) is
a continuous real value, the RC car platform we used unfortunately
only supports three discrete angles---left (-30$^{\circ}$), center
(0$^{\circ}$), and right (+30$^{\circ}$)---as control inputs.
Currently, we approximate the network generated real-valued
angle to the closest one of the three angles, which may
introduce inaccuracy in control.
%% In the future, we plan to use a different (more expensive) RC car
%% platform that can precisely control the car's steering angle.

Other issues we observed in training/testing the network, which
can affect the prediction accuracy of the CNN, are camera and actuator
(motor) control latencies. In DeepPicar's
context, the camera latency is from the time the camera
sensor observes the scene to the time the computer actually reads the
digitized image data. This time can be noticable depending on the camera 
used and the performance of the Pi. Higher camera latency could
negatively affect control performance, because the DNN would analyze
stale scenes. Actuator (motor) control latency also can be a
factor, because moving the steering motor to a desired position takes
time. 
In our platform, the combined latency is measured to be around
50 ms. In other words, in the recorded video and control data, a
control action (left, center, right) taken by the CNN appears to be
applied 50 ms later. Considering that camera's framerate is 30Hz
(33.3 ms/frame), this is about two frames after the control action.
%% We experimentally measured the camera
%% latency and found it to be around 50-100 ms.
If this value is too high, control performance may suffer.
Our initial prototype suffered this problem as the observed camera and
control latency was as high as 300 ms, which negatively affected
control performance.
For reference, the latency of human perception is known to be as fast
as 13 ms~\cite{ThomasBurger2015}. 
% https://www.pubnub.com/blog/2015-02-09-how-fast-is-realtime-human-perception-and-technology/

Despite these issues, the trained models were still able to
achieve a reasonable degree of accuracy, successfully navigating
several different tracks we used for training. In our testing, we found 
that the DeepPicar could remain on a fairly challenging track for over 
10 minutes at a moderate speed (50\% throttle), at which point we stopped
the experiment, and more than one minute at a higher speed (75\%
throttle)~\footnote{A collection of self-driving videos can be found
  at: \fixme{Michael: link your video}
  %% \url{https://photos.app.goo.gl/ce93sU7jPk4ywO8u2}
}. Running at
higher speed is inherently more challenging because the CNN controller
has less time to recover from mistakes (bad predictions).  Also, we
find that the prediction accuracy is significantly affected by the
quality of training data as well as various environmental factors such
as lighting conditions. We plan to investigate more systematic ways
to improve the CNN's prediction accuracies.

We would like to stress, however, that
%% the issues related to the
%% CNN's accuracies have no impact on the \emph{computational 
%% aspects of the system}, and that 
our main focus of this study is not in improving the network accuracy
but in closely replicating the DAVE-2's network architecture and
studying its real-time characteristics, which will be presented in the
subsequent section.
